---
title: "fastHierarchicalReg"
author: "Michael Miller"
date: "2022-11-22"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

fastHierarchicalReg provides functionality for estimating Bayesian hierarchical linear models (HLMs). Bayesian HLMs are very useful in applied statistics because the model learns the prior variance of the model parameters, which functions as a model regularizer, from the data rather than from cross validation (CV). This is critical from an inferential point of view as we can then use the resulting credible intervals to perform inference without having to perform CV for the regularization parameter in addition to calculating bootstrapped confidence intervals.

The primary benchmark used in this package is the programming language Stan, which is a probabilistic programming language for Bayesian models, via the rstan interface. Stan implements the No U-Turn Sampler (NUTS) extension to the Hamiltonian Monte Carlo (HMC) algorithm for sampling from posterior distributions. HMC is state-of-the-art for general posterior sampling, but for HLMs it is possible to design a blocked Gibbs sampler that is faster than the general purpose HMC algorithm implemented in Stan. In particular, by pre-computing the SVD of the design matrix to obtain sufficient statistics for the conditional posterior distributions, we reduce the time complexity of each iteration to $O(p^2)$ in the number of model covariates rather than $O(n^3)$. In addition, by sampling the regression coefficients in one block, we drastically reduce the random walk behavior that Gibbs sampling exhibits for high dimensional problems. The functionality of this package includes

* Parallel computing of the Markov chains using the foreach package
* Fast implementation of the Gibbs sampler using Rcpp and RcppArmadillo
* Calculation of Rhat convergence diagonistics for model parameters


## Installation

```{r, eval = FALSE}
library(devtools)
install_github("mikemiller442/fastHierarchicalReg")
```

## Usage

```{r, message = FALSE}
n <- 10000
numBeta <- 5
betaSD <- 0.75
XSD <- 0.5
errorSD <- 2.0
e <- rnorm(n, mean = 0, sd = errorSD)
beta <- rnorm(numBeta, mean = 0, sd = betaSD*errorSD)
Z <- matrix(NA, nrow = n, ncol = numBeta)
for (i in 1:ncol(Z)) {
  Z[,i] <- rnorm(n, mean = 0, sd = XSD)
}

y <- Z %*% beta + e
output <- y - mean(y)

# Call C++ Function Directly For A Single Chain
res <- linRegGibbs(X = X,
                   testX = testX,
                   Y = Y,
                   testY = testY,
                   numEpochs = numEpochs,
                   regVarPrior = regVarPrior,
                   lambdaSqPrior = lambdaSqPrior)

# Call R Function To Parallel Compute Multiple Chains Using Foreach and Rcpp
res <- linRegGibbsSampler(X = X,
                          testX = testX,
                          Y = output,
                          testY = output,
                          regVarPrior = regVarPrior,
                          lambdaSqPrior = lambdaSqPrior,
                          numEpochs = numEpochs,
                          numDiscard = numDiscard,
                          numChains = numChains)
```





