---
title: "HW3"
author: "Michael Miller"
date: "2022-11-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Calculate Rhat

```{r}
Rhat <- function(samples,numChains) {
  n <- length(samples)/numChains
  si2 <- rep(NA,numChains)
  mui <- rep(NA,numChains)
  for (i in 1:numChains) {
    
    # Calculate chain mean
    chainMean <- samples[(i-1)*n + 1]
    for (j in 2:n) {
      chainMean <- chainMean + samples[(i-1)*n+j]
    }
        
    chainMean <- chainMean/n
    
    # Calculate chain sample variance
    chainVar <- (samples[(i-1)*n + 1] - chainMean)^2
    for (j in 2:n) {
        chainVar <- chainVar + (samples[(i-1)*n+j] - chainMean)^2
    }
    
    chainVar <- chainVar/(n - 1)
    mui[i] <- chainMean
    si2[i] <- chainVar
  }
  
  # Calculate overall mean
  mu <- mui[1]
  for (i in 2:numChains) {
    mu <- mu + mui[i]
  }
  mu <- mu/numChains

  # Calculate average chain sample variance
  s2 <- si2[1]
  for (i in 2:numChains) {
    s2 <- s2 + si2[i]
  }
  s2 <- s2/numChains

  # Calculate scaled between chain sample variance
  B <- (mui[1] - mu)^2
  for (i in 2:numChains) {
    B <- B + (mui[i] - mu)^2
  }
  B <- (n/(numChains - 1))*B

  #Calculate Rhat
  sigmaHat2 <- ((n-1)/n)*s2 + B/n
  Rhat <- sqrt(sigmaHat2/s2)
  return(Rhat)
}
```

### Define Gibbs Sampler

```{r}
library(Rcpp)
library(RcppArmadillo)
library(foreach)

linRegGibbsSampler <- function(X,testX,Y,testY,
                               regVarPrior,lambdaSqPrior,
                               numEpochs,numDiscard,numChains) {
  
  ### Parallel Setup
  n.cores <- parallel::detectCores() - 1
  #create the cluster
  my.cluster <- parallel::makeCluster(n.cores,type = "PSOCK")
  
  #check cluster definition (optional)
  print(my.cluster)
  
  #register it to be used by %dopar%
  doParallel::registerDoParallel(cl = my.cluster)
  
  #check if it is registered (optional)
  foreach::getDoParRegistered()
  
  #how many workers are available? (optional)
  foreach::getDoParWorkers()
  
  ### Run Markov Chains In Parallel
  resMat <- foreach(k = 1:numChains, .combine = "cbind", .packages = c("Rcpp","RcppArmadillo")) %dopar% {
    sourceCpp("./../src/linRegGS.cpp")
    res <- linRegGibbs(X = X,
                       testX = testX,
                       Y = Y,
                       testY = testY,
                       numEpochs = numEpochs,
                       regVarPrior = regVarPrior,
                       lambdaSqPrior = lambdaSqPrior)
    
    postIntercept <- res$intercept[,(numDiscard+2):(numEpochs+1)]
    postBeta <- res$coefBeta[,(numDiscard+2):(numEpochs+1)]
    postLambdaSq <- res$lambdaSq[(numDiscard+2):(numEpochs+1)]
    postLambdaSqScale <- res$lambdaSqScale[(numDiscard+2):(numEpochs+1)]
    postRegVar <- res$regVar[(numDiscard+2):(numEpochs+1)]
    postRegVarScale <- res$regVarScale[(numDiscard+2):(numEpochs+1)]
  
    samples <- rbind(postIntercept,postBeta,postLambdaSq,postLambdaSqScale,
                     postRegVar,postRegVarScale)
    samples
  }
  
  print("Finished sampling chains. Now calculating Rhat.")

  RhatIntercept <- Rhat(resMat[1,],numChains)
  RhatBeta <- rep(NA,ncol(X))
  for (i in 1:ncol(X)) {
    RhatBeta[i] <- Rhat(resMat[1+i,],numChains)
  }
  RhatLambdaSq <- Rhat(resMat[ncol(X)+2,],numChains)
  RhatLambdaSqScale <- Rhat(resMat[ncol(X)+3,],numChains)
  RhatRegVar <- Rhat(resMat[ncol(X)+4,],numChains)
  RhatRegVarScale <- Rhat(resMat[ncol(X)+5,],numChains)
  
  RhatList <- list(RhatIntercept = RhatIntercept,
                   RhatBeta = RhatBeta,
                   RhatLambdaSq = RhatLambdaSq,
                   RhatLambdaSqScale = RhatLambdaSqScale,
                   RhatRegVar = RhatRegVar,
                   RhatRegVarScale = RhatRegVarScale)
  
  postMean <- rowMeans(resMat[,(numDiscard+2):ncol(resMat)])
  postMeanList <- list(intercept = as.numeric(postMean[1]),
                       beta = as.numeric(postMean[2:(ncol(X)+1)]),
                       lambdaSq = as.numeric(postMean[ncol(X)+2]),
                       lambdaSqScale = as.numeric(postMean[ncol(X)+3]),
                       regVar = as.numeric(postMean[ncol(X)+4]),
                       regVarScale = as.numeric(postMean[ncol(X)+5]))
  
  return(list(resMat = resMat,
              postMeanList = postMeanList,
              RhatList = RhatList))
}
```

### Simulate Data 1

```{r}
n <- 10000
numBeta <- 100
betaSD <- 0.75
XSD <- 0.5
errorSD <- 2.0
e <- rnorm(n, mean = 0, sd = errorSD)
beta <- rnorm(numBeta, mean = 0, sd = betaSD*errorSD)
Z <- matrix(NA, nrow = n, ncol = numBeta)
for (i in 1:ncol(Z)) {
  Z[,i] <- rnorm(n, mean = 0, sd = XSD)
}

y <- Z %*% beta + e
output <- y - mean(y)
```

### Run Gibbs Sampler

```{r}
X <- Z
testX <- Z
resp <- output
testResp <- output

numEpochs <- 10000
numDiscard <- 2000
numChains <- 4
lambdaSqPrior <- 1.0
regVarPrior <- 1.0

res <- linRegGibbsSampler(X = X,
                          testX = testX,
                          Y = output,
                          testY = output,
                          regVarPrior = regVarPrior,
                          lambdaSqPrior = lambdaSqPrior,
                          numEpochs = numEpochs,
                          numDiscard = numDiscard,
                          numChains = numChains)
```

### Compare means between model output to true values

```{r}
# Comparing beta
paramBeta <- as.numeric(res$postMeanList$beta)
all.equal(beta,paramBeta)

# Comparing lambda
paramLambda <- sqrt(as.numeric(res$postMeanList$lambdaSq))
all.equal(betaSD,paramLambda)

# Comparing sigma
paramSigma <- sqrt(as.numeric(res$postMeanList$regVar))
all.equal(errorSD,paramSigma)
```

### Simulate Data 2

```{r}
n <- 10000
numBeta <- 5
betaSD <- 0.75
XSD <- 0.5
errorSD <- 2.0
e <- rnorm(n, mean = 0, sd = errorSD)
beta <- rnorm(numBeta, mean = 0, sd = betaSD*errorSD)
Z <- matrix(NA, nrow = n, ncol = numBeta)
for (i in 1:ncol(Z)) {
  Z[,i] <- rnorm(n, mean = 0, sd = XSD)
}

y <- Z %*% beta + e
output <- y - mean(y)
```

### Run Gibbs Sampler

```{r}
X <- Z
testX <- Z
resp <- output
testResp <- output

numEpochs <- 10000
numDiscard <- 2000
numChains <- 4
lambdaSqPrior <- 1.0
regVarPrior <- 1.0

res <- linRegGibbsSampler(X = X,
                          testX = testX,
                          Y = output,
                          testY = output,
                          regVarPrior = regVarPrior,
                          lambdaSqPrior = lambdaSqPrior,
                          numEpochs = numEpochs,
                          numDiscard = numDiscard,
                          numChains = numChains)
```

### Fit the same model in Stan

```{r}
library(rstan)
options(mc.cores = parallel::detectCores())
stanData <- list(N = nrow(X),
                 Y = as.numeric(output),
                 K = ncol(X),
                 X = X)

m1 <- stan(file = "./../m1.stan",
           data = stanData,
           warmup = 1000, 
           iter = 5000,
           control = list(adapt_delta = 0.95),
           chains = 4, 
           cores = 8)

model1 <- data.matrix(m1)
```

### Compare means between model output and Stan

```{r}
# Comparing intercept
stanIntercept <- as.numeric(mean(model1[,1]))
paramIntercept <- as.numeric(res$postMeanList$intercept)
all.equal(stanIntercept,stanIntercept)

# Comparing beta
stanBeta <- as.numeric(colMeans(model1[,2:(ncol(X)+1)]))
paramBeta <- as.numeric(res$postMeanList$beta)
all.equal(stanBeta,paramBeta)

# Comparing lambda
stanLambda <- as.numeric(mean(model1[,ncol(X)+2]))
paramLambda <- sqrt(as.numeric(res$postMeanList$lambdaSq))
all.equal(stanLambda,paramLambda)

# Comparing sigma
stanSigma <- as.numeric(mean(model1[,ncol(X)+3]))
paramSigma <- sqrt(as.numeric(res$postMeanList$regVar))
all.equal(stanSigma,paramSigma)
```

### Compare covariance between model output and Stan

```{r}
# Comparing intercept + beta
stanBetaCov <- as.numeric(cov(model1[,1:(ncol(X)+1)]))
paramBetaCov <- as.numeric(cov(t(res$resMat[1:(ncol(X)+1),])))
all.equal(stanBetaCov,paramBetaCov)

# Comparing lambda
stanLambdaCov <- as.numeric(var(model1[,ncol(X)+2]))
paramLambdaCov <- as.numeric(var(sqrt(res$resMat[ncol(X)+2,])))
all.equal(stanLambdaCov,paramLambdaCov)

# Comparing sigma
stanSigmaCov <- as.numeric(var(model1[,ncol(X)+3]))
paramSigmaCov <- as.numeric(var(sqrt(res$resMat[ncol(X)+4,])))
all.equal(stanSigmaCov,paramSigmaCov)
```

### Rhat Convergence Diagnostic - Check Convergence to 1.0

```{r}
print(unlist(res$RhatList))
all.equal(as.numeric(unlist(res$RhatList)),rep(1.0,length(unlist(res$RhatList))))
```


### Benchmark High Dimensional Example

```{r}
library(bench)

n <- 5000
numBeta <- 500
betaSD <- 0.05
XSD <- 0.5
errorSD <- 2.0
e <- rnorm(n, mean = 0, sd = errorSD)
beta <- rnorm(numBeta, mean = 0, sd = betaSD*errorSD)
Z <- matrix(NA, nrow = n, ncol = numBeta)
for (i in 1:ncol(Z)) {
  Z[,i] <- rnorm(n, mean = 0, sd = XSD)
}

y <- Z %*% beta + e
output <- y - mean(y)

funMod <- function() {
  X <- Z
  testX <- Z
  resp <- output
  testResp <- output
  
  numEpochs <- 4000
  numDiscard <- 2000
  numChains <- 2
  lambdaSqPrior <- 1.0
  regVarPrior <- 1.0
  
  res <- linRegGibbsSampler(X = X,
                            testX = testX,
                            Y = output,
                            testY = output,
                            regVarPrior = regVarPrior,
                            lambdaSqPrior = lambdaSqPrior,
                            numEpochs = numEpochs,
                            numDiscard = numDiscard,
                            numChains = numChains)
  paramIntercept <- as.numeric(res$postMeanList$intercept)
  return(paramIntercept)
}

funStan <- function() {
  options(mc.cores = parallel::detectCores())
  
  X <- Z
  testX <- Z
  stanData <- list(N = nrow(X),
                   Y = as.numeric(output),
                   K = ncol(X),
                   X = X)
  
  m1 <- stan(file = "./../m1.stan",
             data = stanData,
             warmup = 1000, 
             iter = 2000,
             control = list(adapt_delta = 0.95),
             chains = 2,
             cores = 8)
  
  model1 <- data.matrix(m1)
  stanIntercept <- as.numeric(mean(model1[,1]))
  return(stanIntercept)
}

benchMarkRes <- bench::mark(funMod(),
                            funStan(),
                            iterations = 2,
                            check = FALSE)
benchMarkTable <- benchMarkRes[c("expression", "min", "median","mem_alloc","n_gc")]
knitr::kable(benchMarkTable)
```















